---
title: "R Notebook DGE Analysis"
author: Eddie Cho
output: html_notebook
---

#Step 0: load libraries
```{r setup, message=TRUE}
library("DESeq2")
library("BiocParallel")

print("Done. All packages loaded")
```
#Step 1: remove loaded objects then verify current working directory
```{r}
rm(list = ls())
getwd()
```
#Step 2: load FeatureCounts file & Metadata file
```{r}
counts_file<- read.delim("data_set.csv", header = TRUE, row.names = 1, sep =",")

metadata_file <- read.csv("AB_WT42_AB8_Meta.csv")
```

#Step 3-A: verify dimension of featurecounts file & meta data
```{r}
dim(counts_file)
dim(metadata_file)
```
#Step 3-B: display first 6 rows of counts and meta data file
```{r}
head(counts_file)
head(metadata_file)
```
#Step 4: verify if column name of featurecounts file is same as column name of meta data
```{r}
all(rownames(metadata_file)== colnames(counts_file)) 
```
#Step 5: modify row names of meta data
```{r}
rownames(metadata_file) <- metadata_file[,1]

all(rownames(metadata_file)== colnames(counts_file)) 
```

#Step 6: create dds object
```{r}
dds_file <- DESeqDataSetFromMatrix(countData=counts_file, 
                              colData=metadata_file,
                              design= ~condition)
```
#Step 7: perform DESEq2 specific analysis
#estimate gene-wise
```{r}
dds_file_10 <- DESeq(dds_file)
head(dds_file)
```
#Step 8: call results alpha level @ 0.05
```{r}
res0.05 <- results(dds_file_10 , alpha = 0.05)

summary(res0.05)

```
#Step 9: Add organism specific information gene name, etc (Here, Drosophila melanogaster)
```{r}
res0.05$symbol <- mapIds(org.Dm.eg.db,
                     keys=row.names(res0.05), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```
#Step 10: reorder by adjusted p values
```{r}
ord <- order( res0.05$padj )
View(res[ord,])
head(res0.05[ord,])
dim(res0.05[ord,])
```
#Step 11: Export to csv
```{r}
write.csv(res0.05, "DESEQ2_AB_WT42_AB8_DE2_0.05_08_31_23.csv")

print("ALL DESEQ2 PROCESS DONE")
```